---
title: "ISLR Chapter3 "
author: "ACO"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: '6'
  html_document:
    df_print: paged
    toc: true
    toc_depth: '6'
urlcolor: blue
subtitle: 'Linear Regression: Labs'
---

\newpage
```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, warning=FALSE,message=FALSE,attr.source='.numberLines',attr.output='.numberLines', fig.align='center', dpi=350, fig.width=15, fig.height=15)
```

# Linear Regression

## 3.6.1 Libraries

```{r}
library(MASS)
library(ISLR2)
```

## 3.6.2 Simple Linear Regression

Predict `medv` using 12 predictors such as `rm` (average number of rooms per house),
`age` (proportion of owner-occupied units built prior to 1940) and `lstat` (percent
of households with low socioeconomic status).

```{r}
str(Boston, 2)
```

```{r}
lm_fit <- lm(medv ~ lstat , data = Boston)
summary(lm_fit)
```

confidence interval for coeficient estimates

```{r}
confint(lm_fit)
```

confidence interval for predictions

```{r}
predict(
  lm_fit , data.frame(lstat = (c(5, 10, 15))), interval = "confidence")
predict(
  lm_fit , data.frame(lstat = (c(5, 10, 15))), interval = "prediction")
```

```{r}
plot(medv ~ lstat , data = Boston)
abline(lm(medv ~ lstat , data = Boston), col = 'blue')
```

```{r}
par(mfrow = c(2,2))
plot(lm(medv ~ lstat , data = Boston))
par(mfrow = c(1,1))
```

```{r}
par(mfrow = c(2,2))
plot(predict(lm_fit), residuals(lm_fit))
plot(predict(lm_fit), rstudent(lm_fit))
plot(hatvalues(lm_fit))
par(mfrow = c(1,1))
```

```{r}
which.max(hatvalues(lm_fit))
```

## 3.6.3 Multiple Linear Regression

```{r}
lm_fit <- lm(medv ~ lstat + age , data = Boston)
summary(lm_fit)
```

```{r}
lm_fit <- lm(medv ~ ., data = Boston)
summary(lm_fit)
```

Hence R-Squared = `r `summary(lm.fit)$r.sq` and RSE = `summary(lm.fit)$sigma`

```{r}
library(car)
```

```{r}
vif(lm_fit)
```

## 3.6.4 Interaction Terms

```{r}
# interaction, use : or *
lm_fit <- lm(medv ~ lstat:age , data = Boston)
summary(lm_fit)
```

## 3.6.5 Non-linear Transformations of the Predictors

```{r}
lm_fit <- lm(medv ~ lstat , data = Boston)
summary(lm_fit)
```

```{r}
lm_fit2 <- lm(medv ~ lstat + I(lstat^2) , data = Boston)
summary(lm_fit)
```

```{r}
anova(lm_fit, lm_fit2)
```


```{r, eval=F}
lm_fit <- lm(medv ~ lstat + poly(lstat, 2, raw = TRUE) , data = Boston)
summary(lm_fit)
```

## 3.6.6 Qualitative Predictors

```{r}
str(Carseats)
```

```{r}
lm_fit <- lm(
  Sales ~ . + Income:Advertising + Price:Age, data = Carseats)
summary(lm_fit)
```

```{r}
contrasts(Carseats$ShelveLoc)
```


# Applied

## Q8

```{r}
q8_lm = lm(mpg~horsepower, data = Auto)
summary(q8_lm)
```

i. Yes
ii. `r cor(Auto$mpg, Auto$horsepower)`
iii. R squared 60% 
iv 
```{r}
predict(q8_lm, newdata = data.frame(horsepower = 98))
```

```{r}
predict(q8_lm, newdata = data.frame(horsepower = 98), interval = "confidence")
predict(q8_lm, newdata = data.frame(horsepower = 98), interval = "prediction")
```

```{r}
q8_lm = lm(mpg~horsepower, data = Auto)
plot(mpg~horsepower, data = Auto)
abline(q8_lm)
```

```{r}
par(mfrow = c(2,2))
plot(q8_lm)
par(mfrow = c(1,1))
```

## Q9

### a

```{r}
pairs(Auto)
```

### b
```{r}
cor(Auto[, sapply(Auto, FUN = (\(x) is.numeric(x)))])
```

### c
```{r}
q9_model = lm(mpg ~. -name, data = Auto)
summary(q9_model)
```

i. Yes
ii. displacement, weight, year, origin
iii. Most recent cars have higher mpg

### d

```{r}
par(mfrow = c(2,2))
plot(q9_model)
par(mfrow = c(1,1))
```

### e

```{r}
summary(lm(mpg ~ cylinders*displacement, data = Auto))
summary(lm(mpg ~ cylinders:displacement, data = Auto))
```

### f

```{r}
par(mfrow=c(2,2))
q9_lm_f = lm(mpg~weight, data= Auto)
plot(q9_lm_f, 1)

q9_lm_f = lm(mpg~log(weight), data= Auto)
plot(q9_lm_f, 1)

q9_lm_f = lm(mpg~sqrt(weight), data= Auto)
plot(q9_lm_f, 1)

q9_lm_f = lm(mpg~poly(weight, 2, raw = T), data= Auto)
plot(q9_lm_f, 1)
par(mfrow=c(1, 1))
```





























